% created on 2022-08-02
% @author : riverarodrigoa (based on the template of bmazoyer)
\chapter*{Acknowledgements}
\onehalfspacing

Before presenting my work, I would like to wholeheartedly thank everyone that contributed to me becoming a doctor. Whether it involved discussing ideas, polishing arguments, having coffee, playing ping pong, drinking a couple of beers, working out, solving some code or maths problem, gaining confidence, finalizing a paper or delivering the most compelling argument, I want to thank everyone who helped me along the way. As I get to become a doctor, you sure are responsible for it. These words are a token of my everlasting gratitude, and I hope the future is filled with opportunities to work, think, grow, help and laugh with all the people I'm mentioning, and those I may forget. 

First, I would like to thank my two PhD advisors, Lauriane Mouysset and Christopher Costello. Lauriane and I started working together after I finished my masters degree in 2020, before she (immensely) helped me get my PhD funding. Ever since, Lauriane has turned into an amazing mentor, and I hope to soon call her a friend. I want to thank you, Lauriane, for the trust you put in me, your help to grow as a scientist and find interesting questions, your intellectual and emotional support through the difficult process of research. 

I met Chris in 2018, as an intern at the Environmental Markets Lab (emLab), where Chris and his team made my dream to come live in California and work with amazing scientists real. As time went by, Chris went on to become one of my mentors and friends, and welcomed me again and again in Santa Barbara, where I have spent almost three years. Chris, I want to thank you for your vision, that keen eye you have to find interesting ideas and make people feel smart and valuable. I also want to thank you - and your family - for making me feel home in California, for making me a part of your community.

I would like to particularly thank Nicolas Quérou and Agnès Tomini for being the \textit{rapporteurs} of this manuscript, and Martin Quaas, Mireille Chiroleu Assouline, Charles Figuières and Jim Sanchirico, for agreeing to be part of my PhD jury. I am deeply honored that you took part in this jury, as our interactions have been instrumental in the PhD process. Whether I took some of your classes, met you in seminars and conferences, read your insightful articles, your contributions to the field of environmental and resource economics have shaped my understanding, spurred my curiosity, and fueled my work. Thank you!

I am also grateful to the members of my comité de suivi, Maia David, Rodolphe Sabatier and Jean-Luc Dupuy, who have helped me develop my research, channel my enthusiasm and shape my dissertation. Thank you!


My PhD is split between two fabulous research institutions, that make the perfect combination.
First, I want to thank all my colleagues and friends at CIRED. CIRED cultivates an atmosphere of intellectual rigor, cutting-edge research, and convivial, almost family-like environment. The list may not be exhaustive, but here goes. 

I want to thank Romain, for being one of the smartest and funniest people I've shared a desk and conferences with, for your immense help with my projects, and for being real bad at paper toss; Thibault for "créer le commun", for being a dedicated friend, power horse and tremendous DJ; Baptiste, for being the only PSG supporter in that open space, and a brilliant critical mind;  Auriane, for the joy and energy you infused at CIRED, the ping pong trash talk and much more; Lucas, for your sour humour, the endearing support and constructive talks; Mai-Thi, for sharing the good vibes and cultivating a joyful atmosphere; Adrien D., for always questioning what we do; Valentin, for being a source of inspiration in terms of research standards; Yann, for our team and your legendary ping-pong undertables; Charles Claron for teaching me jazz, and opening up my thinking horizons; Pierre Yves (le petit ange), for the intellectual support, and the ping pong beat ups; Charles Collet, for the fun, the econometrics and the chess; Luc, for the ping pong as well, and the opening up of my intellectual perspectives to different horizons; Dimitri, for the cigarette breaks and the progress with \textit{Fences}; Louis, for being the research gold standard and a terrific Bologna tour guide; Soline, for being a good ear, a power through colleague; Morgane, for your soothing perspective on life, your research apetite and your humor; Clément, for your intellectual depth and your fine mustache; Mehdi, for your frienship, energy and motivation; Mathilde for fun times in Rimini and fruitful biodiversity ideas; Bertille, for the beers and help in game theory;  Lucas C, for your talents in horticulture; Bruno, for your support. The faculty at CIRED have also been terrific : thank you Philippe for your ping-pong lessons and care towards the PhD cohort; Adrien F. for your support with the job market; Aurélie, for the support and the dancing; Tarik, for the help with articulating papers; Antoine, for your rigor and interest; Laurent, for the chess and the mathematical help. 
\\
Even though they have already been named, I want to extend a warm thank you to my  PhD seminar co-organizers: Baptiste, Thibault, Clément and Bertille. \\
Finally, I want to thank Carine, Jessy and Valérie for your great administrative help, Estelle for your friendship and making "Y'a pas de forêt sans feux" possible. To everyone I have not mentionned at CIRED, I apologize, and extend the warmest thank you. 


As I have previously stated, I have been lucky enough to be part of two amazing institutions. I first joined emLab when it was in its infancy, and I've been amazed by its culture, way of actually implementing science with a lot of different partners, and convivial work environment. I have met amazing people along the way. I want to thank Julia, my co-author, for bringing me on one of the chapters of this dissertation and becoming a good friend; Kyle, Andrew, Kelsey and Olivier, for their comments, help and letting me learn by doing; members of the Costello Lab, for providing neverending food for thought and productive feedback; Vincent, for Quebec music and the shaping of my final chapter; Jacob, for the wildfire talks and teaching me basketball; Roberto, for our biodiversity talks, and Flavio, for our long discussions, and both for sunday movie night and making me part of the Sueno community, along with Dario and Camilo; Tracey, for being an awesome co-author on seafood and friend in life; Michaela, Heather and Jen for your help and support; and Erin, Lauren, Danielle and Sarah (and Tracey, of course), the brilliant Breaktime Crew. To anyone I may have forgotten at emLab, I apologize, and thank you from the bottom of my heart. I also want to thank my other SB friends, who also made my stay very difficult to end : the Isla Vista Chess Club, Eliott, Marie, Ian, Dan, Phoebe, my housemates Seth and Ferdi, my neighbors Danny, Andrew and Max, my friends in San Francisco, Luis and Joey. 

I have been lucky enough to be surrounded with my long time friends, who listened to me blabing about wildfires, totoaba and linear algebra. Thank you Clara, for being my best friend and housemate along this journey; Leo for our never ending econ talks and travels across the globe; Minha for always opening your home, providing clarity and careful attention;  Jeanne, for being so relentlessly positive, joyful and solutions oriented; Kinan, for the board games and the emotion talks; Joachim and Camille, for traveling the world together and listening all along; Thomas, for the inspiration from your own academic and personal research; Gaspard, for inspiring me to work on nature; Juliette, for your help with competition law and your trust in the future; Louis-Victor, for your positivity and your gear in every situation; Jeremy, for the inspiration from your ambitious projects; Franka, for a:ways being so joyful and true; Alexis, for navigating through life and seemlessly switching directions; Lazare, for being the Nick to our home for 2 years; Ambre, for being so energetic;  Morgane, for your kindness and energy to see the world; Malena, for being slightly crazy, the kindest and supportive ; Robinson, for always being there, stylish, guiding and listening; Maria, for your energy and optimism ; Paul, for exploring the world and still choosing to undertake a PhD;  Augustin A. for your relentless effort to becoming a better version of yourself; Augustin L. for being the next climate grey eminence and the long nights across Paris; Elo\"ise, for your creativity and la GymSu; Clara B., for the fun times in New York; Rico, for dreaming of other horizons together while undertaking a PhD; Julie for the Rimini fun and the EAERE related (more \textit{and} less) fun; Alexandre, for the Platt anecodtes and the endless chess; Tom, for the long nights talking shop at the Bailly; Le\"ila, for your postcards, pep talks and views on everything.

Finally, I need to thank my family and even more than safe, happy places across the globe. Thank you David, Kate, Milla and Alice, for being family for ever, supportive, fun, open-minded and chatty, for welcoming me all this time in New York. Thank you Mambo, Julie and Mateo, for opening your family to me and making California home. Thank you Papi and Mamie for believing in me and wishing success, thank you Mamé, I know you are proud, and thank you Grand Papa, I can say I wish you'd seen that one. Thank you Paul for the torture, and getting me back in shape. Thank you Papa and Maman, or Maman and Papa, for bringing me up, for providing me with everything in life, and for pushing me to pursue great things, for supporting me financially, emotionnally, always and everywhere, thank you. Thank you to my sisters, Julia-Kim and Suzanne, for always believing their brother was smart even though I was pretentious, I have now read enough studies that it is no joke, and to my youngest sibbling, Bob, for being the joyful furrball you are. 

I know this acknowledgements section is already too long, so it's time to conclude. Thank you to everyone involved in this adventure, I can't say it enough, and although I did this for me, I did it because of you, with you as fuel, and I'm immensely grateful for this experience. On to the next one!

%\begin{itemize}
%\item Encadrement : Lauriane, Chris
%\item Jury members : Mireille, Martin, Nicolas, Agnes, Charles, Jim
%\item CIRED : 
%\begin{itemize}
%\item Jeunes : Romain, Thibault, Lucas, Baptiste, Valentin, Auriane, Mai Thi, Charles x2, PY le petit Ange, Luc, Louis, %Soline, Morgane, Clément, Bertille, \textit{Yann, Dimitri}
%\item Titus : Romain, Aurélie, Tarik, Philippe, Antoine, Laurent
%\item Support: Estelle, Carine, Valérie
%\end{itemize}
%\item Emlab : 
%\begin{itemize}
%\item Jeunes : Costello Lab, Jacob, Vincent, Julia, Roberto, Flavio, Dario, Camilo
%\item Life : Seth and Ferdi, Ian
%\item Work: Tracey, Heather, Lauren, Sarah, Danielle, BreakTime Crew
%\end{itemize}
%\item Mes amis, mon équipe : Léo, Jeanne, Minha, Clara, Juliette, LV, Joco, Cam, Charlotte, Thomas; Alex, Morg, Rob, Malé, Maria, Laz, Ambre
%\item Mes autres sangs: Augustin, Louis, Aymeric, Clara à New York, Eloise
%\item Mon père et ma mère, mes frères et soeurs, David, Kate, Alice et Mila pour leur accueil à NY aussi, Mambo et Julie pour m'avoir fait une maison en Californie, mes grands parents, mon cousin Paul pour m'avoir martyrisé, la GymSu en entier, Julie à Louvain; 
%\item Merci du fond du coeur, en entier, pour tout et plus, pour une thèse qui pris 3 ans, deux continents, tant d'amis, de bières, de bêtises et de fêtes. 
%
%\end{itemize}
