---
title: "2_Main model"
author: "Simon"
date: "2024-01-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, 
                      warning = F)
options(scipen=999)

rm(list = ls())

library(tidyverse)
library("ggsci")
library(here)
library(knitr)
library(chromote)
library(stargazer)
library(parallel)
library(webshot2)
library(RColorBrewer)
library(writexl)
library(extrafont)
library(berryFunctions)
library(patchwork)

if(!(c('Fishlife')  %in% installed.packages())){
  devtools::install_github("james-thorson/FishLife")
}
#if(!(c('tinytable') %in% installed.packages())){
#  devtools::install_github("vincentarelbundock/tinytable")
#}

library(tinytable)
#library(FishLife)
library(kableExtra)

Latex = T
font_ = 'roboto'
color_growth = 'black'
color_vertical_monopoly = '#B22222'
color_bertrand = '#1f78b4'
color_cournot = '#248721'
palette_ = c(color_growth, color_vertical_monopoly, color_bertrand, color_cournot)
```

# I. Load data

In this section, I load data for the baseline calibration using the linear and linear quadratic cost models. 


```{r, Load Params}
calibration = read.csv(here("data", 'outputs', "calibration_params.csv"))

for(i in 1:nrow(calibration)){
  assign(calibration$params[i], calibration$value[i])
}  


# Build the 2nd order parameters from parameter list
  # Differentiate parameters for Wild and Farmed inverse demands
alpha_w = alpha
alpha_f = alpha
beta_w = beta
beta_f = beta
  # Demand functions parameters
e   = gamma/(beta_w*beta_f - (gamma^2))
a_f = (alpha_f*beta_w - alpha_w*gamma)/(beta_w*beta_f - (gamma^2))
a_w = (alpha_w*beta_f - alpha_f*gamma)/(beta_w*beta_f - (gamma^2))
b_f = beta_f/(beta_w*beta_f - (gamma^2))
b_w = beta_w/(beta_w*beta_f - (gamma^2))

calibration %>% 
  select(-X)%>%
  mutate(value = round(value,2))  %>%
  kbl() %>%
  kable_styling()

avg_toto_weight = 25.7

#W_lq = read.csv(here('data', 'inputs', 'W_lq.csv'))
W_lq_new = read.csv(here('data','inputs', 'W_lq_new.csv'))
W_mid = W_high

rm(for_display)
```

## II. Define functions
Define function where variable is either `x` i.e population stock, or `s` i.e price paid to poachers. All the parameters take default values specified in the global environment.  

* `growth(x, ...)` : logistic growth function, yields growth of population (in metric tons) 

* `monop_harvest(x, ...)`: harvest (in metric tons) when trader is a monopolist

* `monop_harvest_lq(x, ...)`: harvest (in metric tons) when trader is a monopolist and cost structure is linear quadratic

* `cournot_harvest(x, ...)`, `bertrand_harvest(x, ...)`: harvest (in metric tons) when trader and farmer compete in *Cournot*, i.e, set quantities strategically, and *Bertrand*, i.e, set prices strategically. 

* `cournot_harvest_lq(x, ...)` : harvest  (in metric tons) when trader and farmer compete in *Cournot* i.e, set quantities strategically and cost structure is linear quadratic

* `price_poachers_cournot(x, ...)`, `price_poachers_bertrand(x, ...)`: price paid to poachers (in USD/metric tons) when trader and farmer compete in *Cournot* and *Bertrand*

* `price_poachers_bertrand(x, ...)`: price paid to poachers (in USD/metric tons) when trader and farmer compete in *Bertrand*  and cost structure is linear quadratic

* `bertrand_harvest_lq(x,...)` is set replacing the price that clears the primary market in the harvest function of fishermen. 


* `cournot_farmed(s, ...)`, `bertrand_farmed(s, ...)`: quantity farmed (in metric tons) when trader and farmer compete in *Cournot* and *Bertrand*


```{r}

growth <- function(x, k. = k, r. = r) {
#' Logistic Growth Model Function
#'
#' This function calculates the logistic growth at a given point based on the current value, carrying capacity, and growth rate.
#'
#' @param x Numeric, current population or value at time t.
#' @param k Numeric, carrying capacity of the environment.
#' @param r Numeric, inggtrinsic growth rate of the population.
#'
#' @return Numeric, the calculated growth value at time t.
#' @examples
#' growth(50, 100, 0.02)
#' @export
  y <- r. * x * (1 - x / k.)
  return(y)
}


# II. Monopoly
monop_harvest = function(x, 
                         alpha. = alpha, 
                         c. = c, 
                         sigma. = sigma, 
                         beta. = beta, 
                         W. = W_mid){
#' Monopoly Harvest Function - Quadratic cost model
#'
#' Calculates the harvest level in a vertical monopolistic setting based on a model of quadratic fishery costs in
#' costs (W), transaction costs (c), demand parameters (intercept alpha and sensitivity beta)
#'
#' @param x Numeric, population level at time t
#' @param alpha Numeric, demand intercept in USD
#' @param c Numeric, cost associated with trading
#' @param sigma Numeric, species catchability
#' @param beta Numeric, demand sensisitivity in USD/metric ton
#' @param W Numeric, fishery cost parameter
#'
#' @return Numeric, the calculated monopoly harvest output based on the input parameters.
#' @examples
#' monop_harvest(100, 0.05, 0.01, 0.02, 0.03, 50)
#' @export  

 y = ((alpha. - c.)*sigma.^2 * x^2) / (2*beta.* sigma.^2 * x^2 + 2*W.)
 return(y)
}

monop_harvest_lq = function(x,
                            alpha. = alpha, 
                            c. = c, 
                            sigma. = sigma, 
                            beta. = beta,
                            W1. = W1,
                            W2. = W2){
#' Monopoly Harvest Function - Linear quadratic cost model 
#'
#' Calculates the harvest level in a vertical monopolistic setting based on a model of linear quadratic fishery costs in
#' costs (W), transaction costs (c), demand parameters (intercept alpha and sensitivity beta)
#'
#' @param x Numeric, population level at time t
#' @param alpha Numeric, demand intercept in USD
#' @param c Numeric, cost associated with trading
#' @param sigma Numeric, species catchability
#' @param beta Numeric, demand sensisitivity in USD/metric ton
#' @param W1 Numeric, linear fishery cost parameter, in USD/effort
#' @param W2 Numeric, quadratic fishery effort cost parameter, in USD/effort
#'
#' @return Numeric, the calculated monopoly harvest output based on the input parameters.
#' @examples
#' monop_harvest(100, 0.05, 0.01, 0.02, 0.03, 50)
#' @export  
  

  y = (sigma.^2 * x^2 * (alpha. - c.) - W1.*sigma.*x)/(2*(beta.*sigma.^2 * x^2 + W2.))
  return(y)
}


# III. Cournot

cournot_harvest = function(x, 
                             alpha_f. = alpha_f, 
                             alpha_w. = alpha_w, 
                             beta_f. = beta_f, 
                             beta_w. = beta_w, 
                             gamma. = gamma, 
                             sigma. = sigma, 
                             c. = c, 
                             v. = v, 
                             W. = W_mid){
#' Cournot Competition Harvest Function - Quadratic cost model 
#'
#' Computes the harvest in a Cournot competition setting, considering
#' multiple economic factors like production costs, market demand elasticity, and strategic
#' interaction parameters between firms.
#'
#' @param x Numeric, population level at time t
#' @param alpha_f Numeric, demand intercept in USD for farmed product
#' @param alpha_w Numeric, demand intercept in USD for wild product
#' @param beta_f Numeric, demand sensitivity in USD/metric ton for farmed product
#' @param beta_w Numeric, demand sensitivity in USD/metric ton for wild product
#' @param gamma Numeric, demand sensitivity in USD/metric ton for cross product
#' @param sigma Numeric, catchability
#' @param c Numeric, transaction cost for trader
#' @param v Numeric, marginal cost of production for aquaculture
#' @param W Numeric, fishery effort cost in a quadratic model
#'
#' @return Numeric, the calculated output level based on Cournot competition theory adapted to specific market and firm conditions.

  
  y = ((sigma.^2)*(x^2)*(2*beta_f.*(alpha_w. - c.) - gamma.*(alpha_f. - v.)))/(4*beta_f.*W. + (sigma.^2)*(x^2)*(4*beta_f.*beta_w. - (gamma.^2)))
  return(y)
}

cournot_harvest_lq = function(x, 
                              alpha_f. = alpha_f, 
                              alpha_w. = alpha_w, 
                              beta_f. = beta_f, 
                              beta_w. = beta_w, 
                              gamma. = gamma, 
                              sigma. = sigma, 
                              c. = c, 
                              v. = v, 
                              W1. = W1, 
                              W2. = W2){
#' Cournot Competition Harvest Function - Linear quadratic model 
#'
#' Computes the harvest in a Cournot competition setting, considering
#' multiple economic factors like production costs, market demand elasticity, and strategic
#' interaction parameters between firms.
#'
#' @param x Numeric, population level at time t
#' @param alpha_f Numeric, demand intercept in USD for farmed product
#' @param alpha_w Numeric, demand intercept in USD for wild product
#' @param beta_f Numeric, demand sensitivity in USD/metric ton for farmed product
#' @param beta_w Numeric, demand sensitivity in USD/metric ton for wild product
#' @param gamma Numeric, demand sensitivity in USD/metric ton for cross product
#' @param sigma Numeric, catchability
#' @param c Numeric, transaction cost for trader
#' @param v Numeric, marginal cost of production for aquaculture
#' @param W1 Numeric, linear fishery effort cost in a linear quadratic model
#' @param W2 Numeric, quadratic fishery effort cost in a linear quadratic model
#'
#' @return Numeric, the calculated output level based on Cournot competition theory adapted to specific market and firm conditions.

 y = ((2*beta_f. * (alpha_w. - c.) - gamma. * (alpha_w. -v.))*sigma.^2 * x^2 - 2*beta_f.*W1.*sigma.*x)/(4*beta_f.*W2. + (4*beta_f.*beta_w. - gamma.^2)*sigma.^2 * x^2)
 return(y)
}

price_poachers_cournot = function(x,
                                  alpha_f. = alpha_f, 
                                  alpha_w. = alpha_w, 
                                  beta_f. = beta_f, 
                                  beta_w. = beta_w, 
                                  gamma. = gamma, 
                                  sigma. = sigma, 
                                  c. = c, 
                                  v. = v, 
                                  W. = W_mid){
#' Cournot Competition Price Paid to Poachers Function - Quadratic model 
#'
#' Calculates the price level in a Cournot competition model considering poaching or competitive undercutting,
#' based on parameters that represent strategic interactions, market variability, and firm-specific cost structures.
#'
#' @param x Numeric, population level at time t
#' @param alpha_f Numeric, demand intercept in USD for farmed product
#' @param alpha_w Numeric, demand intercept in USD for wild product
#' @param beta_f Numeric, demand sensitivity in USD/metric ton for farmed product
#' @param beta_w Numeric, demand sensitivity in USD/metric ton for wild product
#' @param gamma Numeric, demand sensitivity in USD/metric ton for cross product
#' @param sigma Numeric, catchability
#' @param c Numeric, transaction cost for trader
#' @param v Numeric, marginal cost of production for aquaculture
#' @param W Numeric, fishery effort cost in a linear quadratic model
#'
#' @return Numeric, the calculated price under Cournot competition 
#' @examples
#' price_poachers_cournot(100, 0.05, 0.04, 0.03, 0.02, 0.01, 0.02, 0.01, 0.015, 50)
#' @export
  y = (2*W.*(2*beta_f.*(alpha_w. - c.) - gamma.*(alpha_f. - v.)))/((sigma.^2)*(x^2)*(4*beta_f.*beta_w. - gamma.^2) + 4*beta_f.*W.)
  return(y)
}

cournot_farmed = function(s, 
                          alpha_f. = alpha_f, 
                          alpha_w. = alpha_w, 
                          beta_f. = beta_f, 
                          beta_w. = beta_w, 
                          gamma. = gamma, 
                          sigma. = sigma, 
                          c. = c, 
                          v. = v){
#' Cournot Competition Model for Farmed Goods
#'
#' Computes the output level in a Cournot competition setting for farmed goods,
#' incorporating market dynamics, competitive interactions, and firm-specific
#' cost structures. The function models the strategic interactions between a
#' firm and its competitor in a market where goods are farmed.
#'
#' @param s Numeric, price paid to poachers in USD
#' @param alpha_f Numeric, demand intercept for farmed product, in USD
#' @param alpha_w Numeric, demand intercept for wild product, in USD
#' @param beta_f Numeric, demand sensitivity for farmed product, in USD/metric ton of biomass
#' @param beta_w Numeric, demand sensitivity for wild product, in USD/metric ton of biomass
#' @param gamma Numeric, demand sensitivity in USD/metric ton of biomass for cross product
#' @param sigma Numeric, catchability
#' @param c Numeric, transaction cost for traders
#' @param v Numeric, marginal cost of farming.
#'
#' @return Numeric, the calculated equilibrium output level for the farming sector under Cournot competition, considering strategic interactions and farm-specific subsidies.
#' @export
  y = (2 * beta_w. * (alpha_f. - v.) - gamma.* (alpha_w. - (s + c.)) )/(4*beta_w.*beta_f. - gamma.^2)
  return(y)
}

# IV. Bertrand

price_poachers_bertrand = function(x, 
                                   a_f. = a_f, 
                                   a_w. = a_w, 
                                   b_f. = b_f, 
                                   b_w. = b_w, 
                                   e.  = e, 
                                   c. = c, 
                                   v. = v, 
                                   sigma. = sigma, 
                                   W. = W_mid){
  y = (2*W.*b_w.*( b_f.*(2*a_w. + e.*v.) + c.*((e.^2) - 2*b_f.*b_w.) + e.*a_f.))/((sigma.^2)*(x^2)*(4*b_f.*b_w. - (e.^2)) + 2*W.*b_w.*(2*b_w.*b_f. - (e.^2)))
  return(y)
}

bertrand_harvest = function(x,
                              sigma. = sigma, 
                              a_f. = a_f, 
                              a_w. = a_w, 
                              b_f. = b_f, 
                              b_w. = b_w, 
                              e. = e, 
                              v. = v, 
                              c. = c, 
                              W. = W_mid){
  y = ((sigma.^2)*(x^2)*b_w.*(b_f.*(2*a_w.+e.*v.) + c.*((e.^2)- 2*b_f.*b_w.) + e.*a_f.))/((sigma.^2)*(x^2)*(4*b_w.*b_f. - (e.^2)) + 2*W.*b_w.*(2*b_f.*b_w. - (e.^2)))
  return(y)
}

bertrand_harvest_lq = function(x,
                               sigma. = sigma, 
                               a_f. = a_f, 
                               a_w. = a_w, 
                               b_f. = b_f, 
                               b_w. = b_w, 
                               e. = e, 
                               v. = v, 
                               c. = c, 
                               W1. = W1, 
                               W2. = W2){
  y1 = b_w.*(sigma.^2 * x^2 *((2*a_w. +e.*v.)*b_f. + e.*a_f.) + (c. + W1. * sigma. * b_w. * x)*(e.^2 - 2*b_w. *b_f.))
  y2 = 2*W2.*b_w.*(2*b_f.* b_w. - e.^2) + (4*b_f.*b_w. - e.^2)*sigma.^2*x^2
  z = y1/y2
  return(z)
}

price_poacher_bertrand_lq = function(x,
                                     sigma. = sigma, 
                                     a_f. = a_f, 
                                     a_w. = a_w, 
                                     b_f. = b_f, 
                                     b_w. = b_w, 
                                     e. = e, 
                                     v. = v, 
                                     c. = c, 
                                     W1. = W1, 
                                     W2. = W2){
  y = 2*b_w.*W2.*((2*a_w. + e.*v.)*b_f. + e.*a_f. + c.*(e.^2 - 2* b_f.*b_w.)) + W1.*sigma.*x*(4*b_f.*b_w. - e.^2)
  z = sigma.^2 * x^2 * (4*b_w.*b_f. - e.^2) + 2*b_w.*W2.*(2*b_f.*b_w - e.^2)
  result = y/z
  return(result)
}

composite_bertrand_lq = function(x,
                               sigma. = sigma, 
                               a_f. = a_f, 
                               a_w. = a_w, 
                               b_f. = b_f, 
                               b_w. = b_w, 
                               e. = e, 
                               v. = v, 
                               c. = c, 
                               W1. = W1, 
                               W2. = W2){
  y = b_w. * ((2*a_w. +e.*v.)*b_f. + e.*a_f. + (price_poacher_bertrand_lq(x)+c.)*(e.^2 - 2*b_w.*b_f))/(4*b_f.*b_w. - e.^2)
  return(y)
}


bertrand_farmed = function(s, 
                             a_f. = a_f, 
                             a_w. = a_w, 
                             b_f. = b_f, 
                             b_w. = b_w, 
                             e. = e, 
                             c. = c, 
                             v. = v){
  y = b_f. *(2*b_w.*a_f. + v.*(e.^2 - 2*b_w.*b_f.) + e.*(a_w. + (s+c.)*b_w.))/(4*b_f.*b_w. - e.^2)
  return(y)
}

```

## III. Generate results

### i. No competition, just traders
```{r, data & analysis}
# Load parameters
calibration = read.csv(here("data", 'outputs', "calibration_params.csv"))

for(i in 1:nrow(calibration)){
  assign(calibration$params[i], calibration$value[i])
}  


# Build the 2nd order parameters from parameter list
  # Differentiate parameters for Wild and Farmed inverse demands
alpha_w = alpha
alpha_f = alpha
beta_w = beta
beta_f = beta
  # Demand functions parameters
e   = gamma/(beta_w*beta_f - (gamma^2))
a_f = (alpha_f*beta_w - alpha_w*gamma)/(beta_w*beta_f - (gamma^2))
a_w = (alpha_w*beta_f - alpha_f*gamma)/(beta_w*beta_f - (gamma^2))
b_f = beta_f/(beta_w*beta_f - (gamma^2))
b_w = beta_w/(beta_w*beta_f - (gamma^2))

W = W_high
# set data : population
data = data.frame(x = seq(1,k))
x_ = seq(1,k)


data = data.frame(x = seq(1,k))%>%
  mutate("Monopoly" = monop_harvest_lq(x_),
         "Quantity adjustment" = cournot_harvest_lq(x_),
         "Price setting" = composite_bertrand_lq(x_))

data %>%
  mutate(Growth = growth(x))%>%
  pivot_longer(cols = -c('x'), # Set to long format for ggplot
              names_to = 'name', 
              values_to = 'values')%>%
  mutate(values = ifelse(values<0,0,values))%>%
  ggplot(aes(x = x, y = values, color = name))+
  geom_line(linewidth = 1)+
  scale_color_manual(values = c("Growth" = "black",
                                "5 trading groups" = "green2",
                                "2 trading groups" = 'green3',
                                "Monopoly" = "#248721",
                                "Quantity adjustment" = "#1f78b4",
                                "Price setting" = "#B22222"))+
  theme_bw()+
  theme(legend.position = 'bottom',
        strip.text.x  = element_text(family = font_, size= 50),
        axis.title.x  = element_text(family = font_, size= 30),
        axis.text = element_text(family = font_),
        legend.text = element_text(family = font_, size = 20), 
        legend.title = element_text(family = font_),
        strip.text.y = element_text(family = font_),
        axis.title.y  = element_text(family = font_, size = 30),
        axis.text.y = element_text(family = font_))+
  labs(color = ' ')+
  xlab('Population (in mt)')+
  ylab('Harvest (in mt)')+  
  guides(
    color = guide_legend(
      title = " ",  # Title for the first column
      nrow = 1
    ))
ggsave(here("C:/Users/jean/OneDrive/Desktop/Projects/PhDissertation/illustrations_for_defense/full.jpg"), width = 10, height = 10, units = "cm")

```